Learning Auditory Similarities Between Instrumental Playing Techniques

### 2020-05-19 paper/paper.tex
cite jansen 2018 icassp

Let $P$ be the number of scattering paths in the operator $\mathbf{S}$. The large-margin nearest neighbor (LMNN) algorithm learns a matrix $\mathbf{L}$ with $P$ rows and $P$ columns by minimizing an error function of the form:
\begin{equation}
\mathcal{E}(\mathbf{L}) = \frac{1}{2} \mathcal{E}_{\textrm{pull}} (\mathbf{L}) + \frac{1}{2} \mathcal{E}_{\textrm{push}} (\mathbf{L})
\end{equation}
Intuitively, the term $\mathcal{E}_{\textrm{pull}}$ tends to shrink local Euclidean neighborhoods in feature space. Crucially, in the case of time--frequency scattering coefficients, these ``Euclidean neighborhoods'' corresponds to Wasserstein neighborhoods in the scalogram domain (see previous subsection).
The definition of $\mathcal{E}_{\textrm{pull}}$ is:
\begin{equation}
\mathcal{E}_{\textrm{pull}} (\mathbf{L}) =
\sum_{\boldsymbol{x}\in\mathcal{X}}
\sum_{\boldsymbol{y}\in\mathcal{Y}_R (\boldsymbol{x})}
\big\Vert
\mathbf{L}\mathbf{\widetilde{S}}\boldsymbol{x} - \mathbf{L}\mathbf{\widetilde{S}}\boldsymbol{y}
\big\Vert^2,
\end{equation}
Note that the error term $\mathcal{E}_\textrm{pull}$ is unsupervised, in the sense that it does not depend on the cluster assignments of $\boldsymbol{x}$ and $\boldsymbol{y}$ in $\mathcal{G}$.
for pairs of samples belonging to the same cluster, while the term $\mathcal{E}_{\textrm{push}}$ tends to ensure that the distance between

\begin{equation}
\mathcal{E}_{\textrm{push}} (\mathbf{L}) =
\sum_{\boldsymbol{x}\in\mathcal{X}}
\sum_{\boldsymbol{y}\in\mathcal{Y}_R (\boldsymbol{x})}
\sum_{\substack{\boldsymbol{z}\in\mathcal{Y}_R (\boldsymbol{x}) \\ \boldsymbol{z}\not\in\mathcal{G}(\boldsymbol{x})}}
\rho
\Big(
1 +
\big\Vert
\mathbf{L}\mathbf{\widetilde{S}}\boldsymbol{x} - \mathbf{L}\mathbf{\widetilde{S}}\boldsymbol{y}
\big\Vert^2
-
\big\Vert
\mathbf{L}\mathbf{\widetilde{S}}\boldsymbol{x} - \mathbf{L}\mathbf{\widetilde{S}}\boldsymbol{z}
\big\Vert^2
\Big),
\end{equation}
where the function $\rho : u \mapsto \max (0, u)$ denotes the activation function of the rectified linear unit (ReLU).

>>>>

Let $P$ be the number of scattering paths in the operator $\mathbf{S}$. The large-margin nearest neighbor (LMNN) algorithm learns a matrix $\mathbf{L}$ with $P$ rows and $P$ columns by minimizing an error function of the form:
\begin{equation}
\mathcal{E}(\mathbf{L}) = \frac{1}{2} \mathcal{E}_{\textrm{pull}} (\mathbf{L}) + \frac{1}{2} \mathcal{E}_{\textrm{push}} (\mathbf{L})
\end{equation}
Intuitively, the term $\mathcal{E}_{\textrm{pull}}$ tends to shrink local Euclidean neighborhoods in feature space. Crucially, in the case of time--frequency scattering coefficients, these Euclidean neighborhoods corresponds to Wasserstein neighborhoods in the scalogram domain (see previous subsection).
The definition of $\mathcal{E}_{\textrm{pull}}$ is:
\begin{equation}
\mathcal{E}_{\textrm{pull}} (\mathbf{L}) =
\sum_{\boldsymbol{x}\in\mathcal{X}}
\sum_{\boldsymbol{y}\in\mathcal{Y}_R (\boldsymbol{x})}
\big\Vert
\mathbf{L}\mathbf{\widetilde{S}}\boldsymbol{x} - \mathbf{L}\mathbf{\widetilde{S}}\boldsymbol{y}
\big\Vert^2,
\end{equation}
Note that the error term $\mathcal{E}_\textrm{pull}$ is unsupervised, in the sense that it does not depend on the cluster assignments of $\boldsymbol{x}$ and $\boldsymbol{y}$ in $\mathcal{G}$.

Conversely, the term $\mathcal{E}_{\textrm{push}}$ tends to penalize small distances between samples that belong to different clusters in $\mathcal{G}$.
More precisely, consider three samples $\boldsymbol{x}$, $\boldsymbol{y}$ and $\boldsymbol{z}$ in $\mathcal{X}$.
Assume that the former belongs to the Euclidean neighborhood of $\boldsymbol{x}$ (i.e., $\boldsymbol{y} \in \mathcal{Y}_R (\boldsymbol{x})$) while the latter belongs to a different cluster as $\boldsymbol{x}$ (i.e., $\boldsymbol{z}\not\in\mathcal{G}(\boldsymbol{x}$)).
The term $\mathcal{E}_{\textrm{push}}$ penalizes $\mathbf{L}$ unless the distance between $\mathbf{L}\mathbf{\widetilde{S}}\boldsymbol{x}$ and $\mathbf{L}\mathbf{\widetilde{S}}\boldsymbol{y}$ is smaller than the distance between $\mathbf{L}\mathbf{\widetilde{S}}\boldsymbol{x}$ and $\mathbf{L}\mathbf{\widetilde{S}}\boldsymbol{z}$ by a margin of at least $1$.
The definition of $\mathcal{E}_{\textrm{push}}$ is:
\begin{equation}
\mathcal{E}_{\textrm{push}} (\mathbf{L}) =
\sum_{\boldsymbol{x}\in\mathcal{X}}
\sum_{\boldsymbol{y}\in\mathcal{Y}_R (\boldsymbol{x})}
\sum_{\boldsymbol{z}\not\in\mathcal{G}(\boldsymbol{x})}
\rho
\Big(
1 +
\big\Vert
\mathbf{L}\mathbf{\widetilde{S}}\boldsymbol{x} - \mathbf{L}\mathbf{\widetilde{S}}\boldsymbol{y}
\big\Vert^2
-
\big\Vert
\mathbf{L}\mathbf{\widetilde{S}}\boldsymbol{x} - \mathbf{L}\mathbf{\widetilde{S}}\boldsymbol{z}
\big\Vert^2
\Big),
\end{equation}
where the function $\rho : u \mapsto \max (0, u)$ denotes the activation function of the rectified linear unit (ReLU), also known as hinge loss.
The cost function described in the equation above is known in deep learning as ``triplet loss'', and has recently been applied to train large-vocabulary audio classifiers in an unsupervised way \cite{jansen2018icassp}.
